# Lynx and hare

## Data

Data was accessed from [github](https://github.com/bblais/Systems-Modeling-Spring-2015-Notebooks/tree/master/data/Lynx%20and%20Hare%20Data) and truncated to the years between 1900 and 1920.

```{r readdata, echo = FALSE}
source(file.path("R", "calc_ccf.R"))
source(file.path("R", "simulate.R"))
data <- read.csv(
  file.path(dir(pattern = "data"), "lynxhare.csv"),
  sep = ";", header = TRUE)
data <- data[data$year %in% 1900:1920, 1:3]
standard <- apply(data[, 2:3], 2,
  function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
```

Each time series is autocorrelated.

```{r autocorrelation, echo = FALSE}
autocorrelation <- sapply(data[, 2:3], acf, na.action = na.pass)
ar1 <- round(c(
  autocorrelation[[1]][-1,1,1][1],
  autocorrelation[[7]][-1,1,1][1]
  ), 2)
ccf(data[, 2], data[, 3], lag.max = 2, plot = FALSE, na.action = na.pass)
```

```{r cor}
marss <- MARSS::MARSS(t(standard),
  model = list(
    "B" = "unconstrained",
    "R" = "zero",
    "U" = "zero",
    "Q" = matrix(list("11", "21", "21", "22")),
    "Z" = "identity"),
  silent = TRUE)
MARSS:::parmat(marss)

MARSS:::parmat(marss)$Q[2,1] /
  (sqrt(MARSS:::parmat(marss)$Q[1,1]) * sqrt(MARSS:::parmat(marss)$Q[2,2]))
# corr/(sqrt(var) * sqrt(var))

# Simulate
set.seed(1)
simulate(MARSS:::parmat(marss))$sim.data
set.seed(1)
MARSS::MARSSsimulate(marss)$sim.data

parccf <- MARSS:::parmat(marss)
stdccf <- calc_ccf(standard, maximumlag = 1)
keep <- stdccf$ccf$acf[2]
parccf$B[c(1,4)] <- sapply(apply(standard, 2,
  forecast::auto.arima, max.p = 1, max.q = 0, seasonal = FALSE),
  "[[", "coef")
parccf$B[c(2:3)] <- 0
# parccf$Q <- matrix(c(1, keep, keep, 1), 2, 2)
parccf$Q <- matrix(c(1, 0, 0, 1), 2, 2)

prod(sqrt(parccf$B[c(1,4)]))

print(parccf)

set.seed(29)
sims <- simulate(parccf, tSteps = NROW(data), nsim = 500)$sim.data
sims2 <- simulate(parccf, tSteps = NROW(data))$sim.data

ccf(sims[1, , 1], sims[2, , 1], plot = FALSE, lag.max = 2)
ccf(sims2[1, , 1], sims2[2, , 1], plot = FALSE, lag.max = 2)
# new <- data.frame(apply(sims, 3, c),
#   "group" = rep(1:NROW(sims[,,1]), NCOL(sims[,,1])),
#   "year" = rep(1:NCOL(sims[,,1]), 2))
# new <- reshape(new, direction = "long",
#     varying = which(grepl("X", colnames(new))), v.names = "X")
```

```{r plot, echo = FALSE}
font <- 0.5
ticks <- c(1.5, 0.25, 0)
line <- 1
lags <- 5
jpeg(file = file.path("results", "rhosimulation_example.jpeg"),
  height = 3, width = 3, units = "in", res = 600)

layout(rbind(c(1, 1), c(2, 3)))
par(
  mar = c(3, 4, 0.5, 3),
  oma = rep(0.25, 4),
  mgp = ticks,
  cex = font,
  cex.axis = font * 1.75,
  tck = -0.2, tcl = -0.2,
  xpd = TRUE
  )
plot(data[, 1], data[, 2] / 1000, type = "l",
  xlab = "year", ylab = "hare (thousands)",
  ylim = c(0, max(data[, 2]) / 1000 * 1.1))
text(x = 1897.25, y = max(data[, 2]) / 1000 * 1.1, labels = "(a)")
par(new = TRUE)
plot(data[, 1], data[, 3] / 1000, type = "l", lty = 2,
  xaxt = "n", yaxt = "n", xlab = "", ylab = "",
  ylim = c(0, max(data[, 3] * 1.1, na.rm = TRUE) / 1000 * 1.1))
legend("topleft", horiz = FALSE,
  legend = c("hare", "lynx"),
  lty = 1:2, bty = "n")
mtext(side = 4, "lynx (thousands)", cex = font, line = line)
axis(side = 4, cex.axis = font * 1.75, mgp = ticks)

par(new = FALSE, xpd = FALSE, mar = c(3, 4, 0.5, 0.1))
aa <- acf(standard[,1], lag.max = lags, ylim = c(-1, 1), xlab = "lag")
lines(aa$lag[,,1] + 0.1,
  acf(standard[,2], na.action = na.pass, lag.max = lags, plot = FALSE)$acf,
  type = "h", lty = 2)
text(x = -1.25, y = 1, labels = "(b)", xpd = TRUE)

par(mar = c(3, 3, 0.5, 1))
aa <- ccf(data[, 2], data[, 3], lag.max = lags, na.action = na.pass,
  plot = FALSE)
plot(aa$lag[,,1], aa$acf[,,1], pch = 19,
  xlab = "lag", ylab = "CCF", ylim = c(-0.45, 0.45))
abline(h = 0)
qi <- qnorm((1 + 0.95)/2)/sqrt(aa$n.used)
abline(h = qi, lty = 2, col = "blue")
abline(h = -qi, lty = 2, col = "blue")

lines(-lags:lags, calc_ccf(standard, maximumlag = lags)$ccf$acf, type = "h")
# lines(-lags:lags + 0.2, calc_ccf(standard[, 2:1], maximumlag = lags)$ccf$acf,
#   lty = 2, type = "h")
text(-1 * lags - 2, 0.45, labels = "(c)", xpd = TRUE)
dev.off()
```

![Visualization of ecological interactions between lynx (Lynx canadensis) and snoeshoe hare (Lepus americanus) from 1900 to 1920. Panel (a) depicts counts (thousands) of hare (solid line) and lynx (dashed line) from 1900 to 1920. Both time series exhibit lag-1 autocorrelation (~0.7; panel b). Cross correlation was only significant at a lag of -1 after an AR(1) model was fit to the hare data to prewhiten the time series (solid vertical lines in panel c). Cross correlation of the raw time series (points in panel c) were not signficant at the 0.05 level. The dashed blue horizontal lines in (b) and (c) are the values beyond which autocorrelations are significantly different from zero.](results/rhosimulation_example.jpeg)

